A cross-cultural universal of human social cognition: Children from 17 diverse communities process gaze in similar ways

### 2024-03-15 paper/paper.tex
add missing site descriptions to supplement

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Unity in diversity: Children from 17 diverse communities process gaze in similar ways},
  pdfauthor={Manuel Bohn1,2, Julia Prein1,2, \ldots\ldots, \& Daniel B. M. Haun2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Unity in diversity: Children from 17 diverse communities process gaze in similar ways}
\author{Manuel Bohn\textsuperscript{1,2}, Julia Prein\textsuperscript{1,2}, \ldots{}\textsuperscript{\ldots{}}, \& Daniel B. M. Haun\textsuperscript{2}}
\date{}

@@

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. Manuel Bohn: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Julia Prein: Conceptualization, Methodology, Writing - Review \& Editing; \ldots{}: Conceptualization, Methodology, Writing - Review \& Editing; Daniel B. M. Haun: Conceptualization, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Manuel Bohn, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute of Psychology in Education, Leuphana University Lüneburg\\\textsuperscript{2} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology}

\abstract{%
Understanding gaze is critical for human social interaction. In a comprehensive cross-cultural study spanning five continents and 17 distinct cultural contexts, we examined the development of gaze understanding among preschoolers using a single reliable task tailored to each cultural setting. Our data provides evidence that children worldwide process gaze in the same way. Key performance signatures of a computational model that sees gaze following as a form of social vector estimation were found in all communities. Additionally, we found a subtle yet cross-culturally consistent relationship between children's environments and their ability to understand gaze. These results highlight the cross-cultural robustness of core social cognitive skills, as well as similarities in the developmental process, suggesting that fundamental aspects of social cognition and interaction emerge and develop in comparable ways globally and hence are a bedrock of human social cognition.
}

@@

Social cognition is a defining aspect of the human species (\emph{1}--\emph{3}). It is supposed to enable unique forms of communication and cooperation that underlie cumulative cultural evolution and the formation of complex societies (\emph{4}--\emph{6}). The eyes are the proverbial window to the mind and the starting point for a majority of social reasoning processes (\emph{7}). Gaze is used to infer the focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation (\emph{8}, \emph{9}). The ability to follow gaze emerges early in childhood (\emph{10}--\emph{12}) and individual differences in children's gaze following ability predict later life outcomes, most notably, later communicative abilities (\emph{13}, \emph{14}). Underlying this narrative is the wide-held assumption that gaze following is fundamental to human social cognition and, therefore, works and develops in the same way across human societies despite substantial variation in developmental contexts. This claim, however, lacks a solid empirical foundation.

We conducted a large-scale cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures and their development. Previous developmental studies focused on the onset of gaze following in infancy (\emph{15}, \emph{16}). The 1377 participants in the study lived in 17 different communities spread over five continents (Fig. \ref{fig:fig1}A, Tab. \ref{tab:tab1}). The countries from which data was contributed to the study represented 46 \% of the world's population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials).

@@

\caption{(A) Data collection sites. Points show approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adpated to each cultural setting. (C) Drawings used as agents across cultural settings. (D) Graphical illustration of the cognitive model. Individuals infer the target of an agent's attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors. Individuals differ in their level of precision (indicated by alpha). For a given level of precision, the further out the target lands, the less precise the model predicts individuals to be. Solid and transparent dots show simualted means and individual data points to illustrate the predicted effect of target position.}\label{fig:fig1}
\end{figure}

\begin{table}

\caption{\label{tab:tab1}Participant demographics.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llllllll}
\toprule
Continent & Country & Community & N(male) & Age (range) & Language & Market integration & Touchscreen\\
\midrule
America & Argentina & Buenos Aires & 105 (53) & 4.72 (3.00 - 6.96) & Spanish (Rioplatense) & high & 0.90\\
 & Mexico & Ocuilan & 127 (63) & 4.96 (2.57 - 6.95) & Spanish (Mexican) & medium & 0.77\\
 & USA & Stanford & 98 (54) & 4.99 (2.52 - 7.90) & English (American) & high & 0.98\\
Africa & Namibia & Hai||om & 60 (38) & 5.85 (2.74 - 8.34) & Hai||om & low & 0.05\\
 &  & Khwe & 59 (24) & 5.84 (3.38 - 8.63) & Khwedam & low & 0.19\\
\addlinespace
 &  & Windhoek & 39 (17) & 5.69 (2.66 - 8.66) & English (Nigerian) & high & 0.95\\
 & Nigeria & Akure & 114 (54) & 5.07 (2.57 - 7.33) & English (Nigerian) & high & 0.91\\
 & Rep. Congo & BaYaka & 29 (13) & 7.80 (3.94 - 10.56) & BaYaka & low & 0.00\\
 &  & Bandongo & 30 (11) & 7.45 (3.50 - 10.95) & Lingala & low & 0.00\\
 & Uganda & Nyabyeya & 125 (62) & 5.94 (2.67 - 8.92) & Bemba & medium & 0.34\\
\addlinespace
 & Zambia & Chimfunshi & 22 (5) & 5.98 (2.88 - 8.00) & Swahili & medium & 0.14\\
Europe & Germany & Leipzig & 100 (48) & 4.88 (2.53 - 6.95) & German & high & 0.89\\
 & UK & Plymouth & 70 (30) & 6.02 (2.38 - 8.94) & English (British) & high & 0.99\\
Asia & China & Beijing & 123 (62) & 5.47 (2.69 - 8.48) & Mandarin & high & 0.95\\
 & India & Pune & 148 (73) & 6.14 (3.06 - 8.83) & English (Indian) / Marathi & high & 0.93\\
\addlinespace
 & Türkiye & Malatya & 85 (40) & 5.02 (2.75 - 7.12) & Turkish & high & 1.00\\
Oceania & New Zealand & Auckland & 43 (19) & 5.14 (2.81 - 8.75) & English (New Zealand) & high & 0.95\\
\bottomrule
\end{tabular}}
\end{table}

We used an animated picture book tablet task in which participants were asked to locate a hidden object based on observing an agent's gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \ref{fig:fig1}B). The key dependent variable was the precision with which children located the agent's focus of attention, that is, the deviation between where the agent looked (where the object was) and where the child thought the agent looked. We adapted visuals and audio instructions specifically to each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task (\emph{17}). Thus, in addition to group-level trends, we were able to investigate individual-level variation.

As the first step, we investigated developmental gains. Across all 17 communities, we found a substantial increase in average levels of precision with age (fixed effect in Bayesian regression model (\emph{18}): \(\beta\) = -0.30, 95\% HDI (-0.40 - -0.21); range of community-level effects in model: \(\beta_{min}\) = -0.06, 95\% HDI (-0.17 - 0.06) to \(\beta_{max}\) = -0.59, 95\% HDI (-0.70 - -0.47). Nevertheless, there were also marked differences between communities (see Fig. \ref{fig:fig2}A). In a cross-validation procedure, in which we trained a regression model on a subset of the data (training data) to later predict the held out data (testing data) (\emph{19}), we found that a model that assumed cross-cultural variation to outperform simple models in 100\% of cases (see Supplemental Material). At first glance, it seems that highly market-integrated communities around the globe showed higher levels of precision compared to less market-integrated communities (see Tab. \ref{tab:tab1}). However, believe that these results are better understood in terms of exposure to technology and we discuss them in more detail below. Furthermore, average differences between communities were small compared to differences between individuals: communities did not form homogeneous clusters but largely overlapping distributions in that some individuals from communities with a lower average level of precision performed better compared to some individuals from a setting with a very high average level of precision. Similarly, in all communities, some 4-year-olds outperformed children two years older than them (see Fig. \ref{fig:fig2}A).

@@

{\centering \includegraphics[width=0.75\linewidth]{../visuals/pvis_pred}

@@

\caption{A) Developmental trajectories by cultural settings. Transparent dots show aggregated data for each individual. The color dentoes the different cultural settings. The developmental trajectories are predicted based on a model of the data aggregated for each participant. B) Performance by screen section and cultural setting. Each bin covers 1/10th of the screen. Points show means and eroor bars 95\% confidence intervals for the data within that bin aggregated across participants.}\label{fig:fig2}
\end{figure}

Consistent developmental gains alone cannot inform us about the cognitive processes children use when locating the agent's focus of attention. Recent computational work modeled gaze following as social vector estimation (\emph{20}). When observing the eyes, onlookers estimate a vector from the center of the eye through the pupil. The focus of attention is the location wherever the estimated vectors from both eyes hit a surface (Fig. \ref{fig:fig1}C). The estimation process is not perfect, but each individual has a systematic level of uncertainty, which is conceptualized as the cause of individual differences. Importantly, this process model predicts a clear performance signature in our cross-cultural gaze following task: Trials in which the agent looks further away from the center (i.e.~to the left or right side of the screen) should result in lower levels of precision, compared to trials in which the agent looks closer to the middle. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual's level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area.

The processing signature was clearly visible across all 17 communities. Precision decreased when the agent looked at locations further away from the center (fixed effect: \(\beta\) = 0.47, 95\% HDI ( - 0.54); range of community-level effects: \(\beta_{min}\) = 0.58, 95\% HDI (0.51 - 0.66) to \(\beta_{max}\) = 0.16, 95\% HDI (-0.01 - 0.34). Visualization of the data showed the predicted u-shaped pattern in all communities (see Fig. \ref{fig:fig2}B). These results indicate a universal cognitive process used by children in all communities. There is, however, an alternative way in which the u = shaped pattern might arise: it would also arise when participants ignored the agent's gaze and instead always selected the middle of the screen (center bias). To rule out this alternative explanation, we directly compared three cognitive models that made different assumptions about how participants' responses were generated: the focal vector-based gaze estimation model described above, a center-bias model where participants always select the center, and a random guessing model where participants select random locations. For every community, we found overwhelming support for the gaze estimation model (min \(BF_{10}\) \textgreater{} 1000 for comparisons with both alternative models). Taken together, children from all 17 communities processed gaze in similar ways.

What is left to explain are the marked community- and individual-level differences. In addition to the gaze-following task, caregivers filled out a small questionnaire about children's access to screen-based technology and the composition of their households. On an individual level, we found that across communities, children with access to touchscreen devices had higher levels of precision (\(\beta\) = , SE = , 95\% CrI = - ). On a community level, we also saw that average performance was lowest in the communities in which touchscreen devices were the least frequent (Fig. \ref{fig:fig3}). Thus, methodological factors likely explained the marked differences between communities.

{\centering \includegraphics[width=0.5\linewidth]{../visuals/pcomtc}

\caption{Correlation between performance in the gaze-following task and the exposure to touchscreens on a cultural setting-level. Realtive imprecision accounts for differences in age between settings in that it corresponds to the random intercept estimate for each setting in a model predicting performance by age.}\label{fig:fig3}
\end{figure}

Based on this, we also computed the number of older children living in the household. We did not pre-register this, but decided to also include this variable because prior work on advanced forms of social cognition has suggested that the presence of older siblings is slightly more predictive of development compared to other aspects of household composition (e.g., \emph{21}, \emph{22}, but see \emph{23}). We think that gaze following improves though practice during social interaction. Particularly supportive, we believe, are situations in which it is difficult to infer the goals and intentions of another person based on other information sources (e.g., language or context). In such situations, children have to pay closer attention to the interaction partner and use their gaze to make such inferences. We are well aware that the numbers in the questionnaire are very far removed from these forms of social interaction. Thus, at best, they are a very coarse approximation. As a consequence, we do not expect there to be a very strong link between the two. We sincerely hope that future work will be able to directly assess children's social interactions.

On an individual level, other factors likely generated individual differences because there was substantial variation even in communities where almost all children had access to touchscreens. Previous work suggested that social cognition develops in social interaction (\emph{24}--\emph{26}), in particular, with older siblings (\emph{21}, \emph{22}). We found tentative support for this idea in the form of a small but consistent effect in the direction that participants living together with more children older than themselves (relative to other children from the same setting) had slightly higher levels of precision (\(\beta\) = , SE = , 95\% CrI = - ).

>>>>

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={A cross-cultural universal of human social cognition: Children from 17 diverse communities process gaze in similar ways},
  pdfauthor={Manuel Bohn (ORCID: 0000-0001-6006-1348)1,2,*, Julia Prein (ORCID: 0000-0002-3154-6167)1,2,*, Agnes Ayikoru3, Florian M. Bednarski (ORCID: 0000-0003-4384-4791)4, Ardain Dzabatou5, Michael C. Frank (ORCID: 0000-0002-7551-4378)6, Annette M. E. Henderson (ORCID: 0000-0003-4384-4791)4, Joan Isabella3, Josefine Kalbitz2, Patricia Kanngiesser (ORCID:0000-0003-1068-3725)7, Dilara Keşşafoğlu (ORCID: 0000-0002-7356-0733)8, Bahar Köymen (ORCID: 0000-0001-5126-8240)9, Maira V. Manrique-Hernandez2, Shirley Magazi (ORCID: 0009-0006-0479-9800)10, Lizbeth Mújica-Manrique2, Julia Ohlendorf2, Damilola Olaoba2, Wesley R. Pieters (ORCID:0000-0002-6152-249X)10, Sarah Pope-Caldwell2, Katie Slocombe (ORCID: 0000-0002-7310-1887)11, Robert Z. Sparks (ORCID: 0000-0001-7545-0522)6, Jahnavi Sunderarajan2, Wilson Vieira2, Zhen Zhang (ORCID: 0000-0001-9300-0920)12, Yufei Zong (ORCID: 0009-0000-5012-0244)12, Roman Stengelin (ORCID: 0000-0003-2212-4613)2,10, \& Daniel B. M. Haun2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{A cross-cultural universal of human social cognition: Children from 17 diverse communities process gaze in similar ways}
\author{Manuel Bohn (ORCID: 0000-0001-6006-1348)\textsuperscript{1,2,*}, Julia Prein (ORCID: 0000-0002-3154-6167)\textsuperscript{1,2,*}, Agnes Ayikoru\textsuperscript{3}, Florian M. Bednarski (ORCID: 0000-0003-4384-4791)\textsuperscript{4}, Ardain Dzabatou\textsuperscript{5}, Michael C. Frank (ORCID: 0000-0002-7551-4378)\textsuperscript{6}, Annette M. E. Henderson (ORCID: 0000-0003-4384-4791)\textsuperscript{4}, Joan Isabella\textsuperscript{3}, Josefine Kalbitz\textsuperscript{2}, Patricia Kanngiesser (ORCID:0000-0003-1068-3725)\textsuperscript{7}, Dilara Keşşafoğlu (ORCID: 0000-0002-7356-0733)\textsuperscript{8}, Bahar Köymen (ORCID: 0000-0001-5126-8240)\textsuperscript{9}, Maira V. Manrique-Hernandez\textsuperscript{2}, Shirley Magazi (ORCID: 0009-0006-0479-9800)\textsuperscript{10}, Lizbeth Mújica-Manrique\textsuperscript{2}, Julia Ohlendorf\textsuperscript{2}, Damilola Olaoba\textsuperscript{2}, Wesley R. Pieters (ORCID:0000-0002-6152-249X)\textsuperscript{10}, Sarah Pope-Caldwell\textsuperscript{2}, Katie Slocombe (ORCID: 0000-0002-7310-1887)\textsuperscript{11}, Robert Z. Sparks (ORCID: 0000-0001-7545-0522)\textsuperscript{6}, Jahnavi Sunderarajan\textsuperscript{2}, Wilson Vieira\textsuperscript{2}, Zhen Zhang (ORCID: 0000-0001-9300-0920)\textsuperscript{12}, Yufei Zong (ORCID: 0009-0000-5012-0244)\textsuperscript{12}, Roman Stengelin (ORCID: 0000-0003-2212-4613)\textsuperscript{2,10}, \& Daniel B. M. Haun\textsuperscript{2}}
\date{}

@@

The authors would like to thank Luke Maurits for statistical advice. Manuel Bohn was supported by a Jacobs Foundation Research Fellowship (2022-1484-00). We are greatful to thank all children and caregivers for participating in the study.

The authors made the following contributions. Manuel Bohn (ORCID: 0000-0001-6006-1348): Conceptualization, Methodology, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Julia Prein (ORCID: 0000-0002-3154-6167): Conceptualization, Methodology, Software, Investigation, Writing - Review \& Editing; Agnes Ayikoru: Investigation; Florian M. Bednarski (ORCID: 0000-0003-4384-4791): Investigation, Writing - Review \& Editing; Ardain Dzabatou: Investigation; Michael C. Frank (ORCID: 0000-0002-7551-4378): Investigation, Writing - Review \& Editing; Annette M. E. Henderson (ORCID: 0000-0003-4384-4791): Investigation, Writing - Review \& Editing; Joan Isabella: Investigation; Josefine Kalbitz: Investigation, Writing - Review \& Editing; Patricia Kanngiesser (ORCID:0000-0003-1068-3725): Investigation, Writing - Review \& Editing; Dilara Keşşafoğlu (ORCID: 0000-0002-7356-0733): Investigation, Writing - Review \& Editing; Bahar Köymen (ORCID: 0000-0001-5126-8240): Investigation, Writing - Review \& Editing; Maira V. Manrique-Hernandez: Investigation; Shirley Magazi (ORCID: 0009-0006-0479-9800): Investigation; Lizbeth Mújica-Manrique: Investigation, Writing - Review \& Editing; Julia Ohlendorf: Investigation; Damilola Olaoba: Investigation; Wesley R. Pieters (ORCID:0000-0002-6152-249X): Investigation, Writing - Review \& Editing; Sarah Pope-Caldwell: Investigation; Katie Slocombe (ORCID: 0000-0002-7310-1887): Investigation; Robert Z. Sparks (ORCID: 0000-0001-7545-0522): Investigation; Jahnavi Sunderarajan: Investigation; Wilson Vieira: Investigation; Zhen Zhang (ORCID: 0000-0001-9300-0920): Investigation, Writing - Review \& Editing; Yufei Zong (ORCID: 0009-0000-5012-0244): Investigation, Writing - Review \& Editing; Roman Stengelin (ORCID: 0000-0003-2212-4613): Conceptualization, Methodology, Investigation, Writing - Review \& Editing; Daniel B. M. Haun: Conceptualization, Funding acquisition, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Manuel Bohn (ORCID: 0000-0001-6006-1348), Universitätsallee 1, 21335 Lüneburg, Germany. E-mail: \href{mailto:manuel.bohn@leuphana.de}{\nolinkurl{manuel.bohn@leuphana.de}}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute of Psychology in Education, Leuphana University Lüneburg\\\textsuperscript{2} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology\\\textsuperscript{3} Budongo Conservation Field Station\\\textsuperscript{4} School of Psychology, University of Auckland\\\textsuperscript{5} Université Marien Ngouabi\\\textsuperscript{6} Department of Psychology, Stanford University\\\textsuperscript{7} School of Psychology, University of Plymouth\\\textsuperscript{8} Department of Psychology, Koç University\\\textsuperscript{9} Division of Psychology, Communication, and Human Neuroscience, University of Manchester\\\textsuperscript{10} Department of Psychology and Social Work, University of Namibia\\\textsuperscript{11} Department of Psychology, University of York\\\textsuperscript{12} CAS Key Laboratory of Behavioral Science, Institute of Psychology, Chinese Academy of Sciences\\\textsuperscript{*} joint first author}

\abstract{%
Theoretical accounts assume that key features of human social cognition are universal. Here we focus on gaze following, the bedrock of social interactions and coordinated activities, to test this claim. In this comprehensive cross-cultural study spanning five continents and 17 distinct cultural communities, we examined the development of gaze following in early childhood. We identified key processing signatures through a computational model that assumes that participants follow an individual's gaze by estimating a vector emanating from the eye-center through the pupil. Using a single reliable touchscreen-based task, we found these signatures in all communities, suggesting that children worldwide processed gaze in highly similar ways. Additionally, we found a cross-culturally consistent relationship between children's exposure to touchscreens and their performance in the task, which also explained absolute differences between communities. These results provide strong evidence for a universal process underlying a foundational socio-cognitive ability in humans that can be reliably inferred even in the presence of cultural variation in overt behavior.
}

@@

Human socio-cognitive skills enable unique forms of communication and cooperation that provide a bedrock for cumulative culture and the formation of complex societies (\emph{1}--\emph{7}). The eyes are the proverbial ``window to the mind'' and eye gaze is essential for many social reasoning processes (\emph{8}, \emph{9}). Others' eye gaze is used to infer their focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation (\emph{10}, \emph{11}). The ability to follow gaze emerges early in childhood (\emph{12}--\emph{15}) and individual differences in children's gaze-following ability predict later life outcomes, most notably, communicative abilities (\emph{16}, \emph{17}). Difficulties with gaze following have been linked to developmental disorders, including Autism (\emph{18}).

There is a widely-held assumption, that, despite substantial variation in developmental contexts, gaze-following works and develops in the same way across human societies (\emph{19}). However, this view is seemingly at odds with the substantial diversity in socio-cognitive development that cross-cultural studies have revealed (\emph{3}, \emph{20}--\emph{23}). One potential source for this paradox lies in the reliance on aggregated measures in cross-cultural studies. Absolute differences in mean performance across communities are interpreted as a signal of different underlying cognitive processes. In this study, we resolve this paradox by instead focusing on processing signatures that can be investigated independently of absolute community-level differences. This allows us to directly evaluate the empirical foundation of claims about universal features of human social cognition. To this end, we conducted a pre-registered, large-scale, cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures. These signatures were derived from a computational model that assumes that participants follow gaze by estimating a vector emanating from the eye center through the pupil (\emph{24}).

The 1377 participants who took part in the study lived in 17 different communities across 14 countries and five continents (Fig. \ref{fig:fig1}A, Tab. \ref{tab:tab1}). These countries represent \textasciitilde46\% of the world's population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials). This diversity allowed us to overcome the common pitfall of cross-cultural studies that largely compared urban communities from the global north to rural communities from the global south (\emph{25}).

@@

\caption{(A) Data collection sites. Points show the approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the location on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adapted to each cultural setting. (C) Drawings used as agents across cultural settings. (D) Graphical illustration of the cognitive model. Individuals infer the target of an agent’s attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors (transparent lines). Individuals differ in their level of precision (indicated by sigma). For a given level of precision, the closer the  target lands to the left or the right, the less precise the model predicts individuals to be. Solid and transparent dots show simulated means and individual data points to illustrate the predicted effect of target position.}\label{fig:fig1}
\end{figure}

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} *Local collaborators and piloting suggested that Nigerian English is suitable for Windhoek as well.}
\end{TableNotes}

\begin{longtable}{m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}m{0.111111111111111\linewidth}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:tab1}Participant demographics.}\\
\toprule
Continent & Country & Community & N(male) & Age (range) & Language & Market integration & Touchscreen\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:tab1} continued}}\\
\toprule
Continent & Country & Community & N(male) & Age (range) & Language & Market integration & Touchscreen\\
\midrule
\endhead
Americas & Argentina & Buenos Aires & 105 (53) & 4.72 (3.00 - 6.96) & Spanish (Rioplatense) & high & 0.90\\
 & México & Ocuilan & 127 (63) & 4.96 (2.57 - 6.95) & Spanish (Mexican) & medium & 0.77\\
 & USA & Stanford & 98 (54) & 4.99 (2.52 - 7.90) & English (American) & high & 0.98\\
Africa & Namibia & Hai||om & 60 (38) & 5.85 (2.74 - 8.34) & Hai||om & low & 0.05\\
 &  & Khwe & 59 (24) & 5.84 (3.38 - 8.63) & Khwedam & low & 0.19\\
 &  & Windhoek & 39 (17) & 5.69 (2.66 - 8.66) & English (Nigerian) & high & 0.95\\
 & Nigeria & Akure & 114 (54) & 5.07 (2.57 - 7.33) & English (Nigerian)* & high & 0.91\\
 & Rep. Congo & BaYaka & 29 (13) & 7.80 (3.94 - 10.56) & BaYaka & low & 0.00\\
 &  & Bandongo & 30 (11) & 7.45 (3.50 - 10.95) & Lingala & low & 0.00\\
 & Uganda & Nyabyeya & 125 (62) & 5.94 (2.67 - 8.92) & Swahili & medium & 0.34\\
 & Zambia & Chimfunshi & 22 (5) & 5.98 (2.88 - 8.00) & Bemba & medium & 0.14\\
Europe & Germany & Leipzig & 100 (48) & 4.88 (2.53 - 6.95) & German & high & 0.89\\
 & UK & Plymouth & 70 (30) & 6.02 (2.38 - 8.94) & English (British) & high & 0.99\\
Asia & China & Beijing & 123 (62) & 5.47 (2.69 - 8.48) & Mandarin & high & 0.95\\
 & India & Pune & 148 (73) & 6.14 (3.06 - 8.83) & English (Indian) / Marathi & high & 0.93\\
 & Türkiye & Malatya & 85 (40) & 5.02 (2.75 - 7.12) & Turkish & high & 1.00\\
Oceania & New Zealand & Auckland & 43 (19) & 5.14 (2.81 - 8.75) & English (New Zealand) & high & 0.95\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

\end{ThreePartTable}
\end{center}

We used an animated picture book tablet task in which participants had to locate a hidden object based on observing an agent's gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \ref{fig:fig1}B). The key dependent variable was the (im)precision with which children located the agent's focus of attention, that is, the deviation between where the agent looked (where the object was) and the child's response. We adapted visuals and audio instructions specifically to each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task in a German sample (\emph{26}).

As the first step, we investigated developmental improvements, that is, how children become more precise at estimating the target location with age. Across all 17 communities, we found a substantial increase in average levels of precision with age (fixed effect in Bayesian regression model (\emph{27}): \(\beta\) = -0.30, 95\% HDI (-0.40 - -0.21); range of community-level (random) effects: \(\beta_{min}\) = -0.06, 95\% HDI (-0.18 - 0.05) to \(\beta_{max}\) = -0.59, 95\% HDI (-0.71 - -0.48).

Nevertheless, there were also marked differences between communities (see Fig. \ref{fig:fig2}A). In a six-fold cross-validation procedure, we trained a regression model on a subset of the data (training data) to later predict the held-out data (testing data) (\emph{28}). We found that a model assuming cross-cultural variation in average performance as well as cross-cultural variation in developmental trajectories outperformed simpler models -- assuming no variation in the shape of developmental trajectories or nor variation between settings at all -- in 98\% of cases (see Supplemental Material). At first glance, it seems that highly market-integrated communities around the globe showed higher levels of precision compared to less market-integrated communities (compare Tab. \ref{tab:tab1} and Fig. \ref{fig:fig2}A). However, numerous alternative groupings (e.g., average levels of education or average household size) are possible and, instead, these results could be best understood in terms of exposure to touch-screen devices; a finding we discuss in more detail below. Importantly, average differences in precision between communities were small compared to differences between individuals: communities did not form homogeneous clusters but largely overlapping distributions in that some individuals from communities with a lower average level of precision performed better compared to some individuals from a setting with a very high average level of precision. Similarly, in all communities, some 4-year-olds outperformed children two years older than them (see Fig. \ref{fig:fig2}A). The lack of adequate individual-level measurement instruments in previous large-scale developmental cross-cultural studies made it impossible to contrast these perspectives. The substantial overlap between communities found here speaks against categorical differences in gaze following and is suggestive of a similar underlying process. However, consistent developmental improvements alone cannot inform us about the cognitive processes children use when locating the agent's focus of attention.

@@

{\centering \includegraphics[width=1\linewidth]{../visuals/pvis_pred}

@@

\caption{A) Developmental trajectory across and B) by community. The developmental trajectories are predicted based on a model of the data aggregated for each participant. C) Performance by screen section across and D) by community. Each bin covers 1/10th of the screen. Points show means, and error bars 95\% confidence intervals for the data within that bin aggregated across participants. Transparent dots in A) and C) show aggregated data for each individual.}\label{fig:fig2}
\end{figure}

Recent computational work modeled gaze following as social vector estimation (\emph{24}). When observing the eyes, onlookers estimate a vector emanating from the center of the eye that runs through the pupil. The focus of attention is the location where the estimated vectors from both eyes hit a surface (Fig. \ref{fig:fig1}D). It is assumed that this estimation process has some uncertainty and that individuals vary in their level of uncertainty. As a consequence, even though individuals use the same general process, they might differ in their absolute levels of precision. Crucially, this process model predicts a clear performance signature in our gaze following task: Trials in which the agent looks further away from the center (i.e.~to the left or right side of the screen) should result in lower levels of precision compared to trials in which the agent looks closer to the middle. This prediction is best understood by considering a similar phenomenon: pointing a torch light to a flat surface. The width of the light beam represents each individual's level of uncertainty in vector estimation. When the torch is directed straight down, the light beam is concentrated in a relatively small area. When the torch is rotated to the side, the light from one half of the cone must travel further than the light from the other half to reach the surface. As a consequence, the light is spread over a wider area (see Fig. \ref{fig:fig1}D).

This processing signature was clearly visible across all 17 communities. Precision decreased when the agent looked at locations further away from the center (fixed effect: \(\beta\) = 0.47, 95\% HDI ( - 0.54); range of community-level effects: \(\beta_{min}\) = 0.58, 95\% HDI (0.51 - 0.66) to \(\beta_{max}\) = 0.16, 95\% HDI (-0.01 - 0.33). Visualization of the data showed the predicted u-shaped pattern in all communities (see Fig. \ref{fig:fig2}B). These results indicate a universal cognitive process used by children in all communities. There are, however, alternative ways in which the u-shaped pattern might arise: if participants ignored the agent's gaze and instead always selected the middle of the screen (center bias) or randomly select locations (random guessing), precision would also decrease when the balloon lands further away from the center. To rule out these alternative explanations, we compared three cognitive models that made different assumptions about how participants' responses were generated: the focal vector-based gaze estimation model described above, a center-bias model where participants always select the center, and a random guessing model where participants select random locations. For every community, we found overwhelming support for the gaze estimation model (min \(BF_{10}\) \textgreater{} 100 000 for comparisons with both alternative models). Taken together, children from all 17 communities processed gaze in similar ways.

Next, we looked at factors that could explain community- and individual-level variation. In addition to the gaze-following task, caregivers responded to a short questionnaire about children's access to screen-based technology and household composition. On an individual level, we found that children with access to touchscreen devices had higher levels of precision (\(\beta\) = -0.14, SE = 0.04, 95\% CrI = -0.21 - -0.07). This effect was consistent across communities in that allowing the effect of access to touchscreens to vary across communities did not improve model fit (see Supplemental Materials). On a community level, we also saw that average performance was lowest in communities in which touchscreen devices were the least frequent (community-level correlation between age-corrected imprecision and proportion of children with access to touchscreens: \emph{r} = -0.90, 95\% CI = -0.96 - -0.74). Thus, familiarity with the device used for data collection likely explains variation between communities. The most likely explanation for this effect is that children with more touchscreen experience were better at task handling and thus more likely to precisely touch the location they inferred the agent to look at.
